\chapter{\textbf{Conclusion}}
The fabricated system run successfully and 20 trials were performed to measure its performance. It showed no errors in those 20 trials which took almost 3 hours to execute. The control system took almost 1 seconds (approximately) to perform necessary commands when signals came from the host server. From analyzed signals the impulse period and the group delay period seemed quite perfect if the Baud Rate (Sampling Frequency) tuning ranges from 11 to 12.6 kHz. In this dissertation, 11.52 kHz Sampling Frequency was used and the power density spectrum of gas sensor was quite fine though the Signal to Noise ratio was a little high. This can be overcome by using a 20 pF ceramic disk capacitor as it can work as a low pass filter. The overall system efficiency was about 95$\%$ , which is quite good for a robust controlling operation.\\
Work With the continuous development in technology and availability of internet it is possible to develop a low-cost smart sensor node which enables things to be connected and monitored easily and corresponding information can be accessible globally. An integrated framework is
developed with the combination of WSN and Local IP. The promising issue of the proposed method is to provide a flexible low-cost solution for integrating Local IP  with home automation or Industrial automation systems. A real time application of WSN for the development of smart home or industry is designed and implemented. The smart environment is created by adapting different types of sensors and actuators which are used to communicate with things through webpage over the internet. The gathered sensor data from different sensor node are displayed in the user dashboard. User can monitor the environment and control the physical objects. Also, to give more intelligence to the home devices, a decision-making algorithm is designed based on http request.
\section{Future Research Directions}
This framework can be extended for the development of building and city automation and Industry 4.0. Also, the WSN based control scheme can be extended for controlling and monitoring of remote devices such as robots, different types of rovers etc. Though there are numerous future scopes of the project, some of this are listed below
The web page further can be developed to provide real time chart for better visualization of sensor node data in an android APK app for the Android users.\\
Voice commands technology can be implemented for controlling Actuators.\\
Adaptive fuzzy logic can be implemented for automatic controlling of AC, water pump and other equipment’s. It can also be used to pre-detect the probability of the fire.Fuzzy logic can be  implemented for the control of environment temperature. The decision-making system can provide 99:2\% accuracy with respect to the MATLAB simulation results.\\
The web page further can be developed to provide real time chart for better visualization of sensor node data. Power consumption of the equipment’s can also be presented in the webpage.\\
Image processing can be implemented to find out known and unknown faces. It can be used as security purposes by detecting faces.
